---
title: "Interactive Visualizations in R"
author: "Jered Ataky"
date: "10/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries

Loading necessary libraries 

```{r warning=FALSE, message=FALSE}
library(shiny)
library(plotly)
library(rsconnect)
```

## Load the Data

```{r}
cdc_data <- read.csv("https://raw.githubusercontent.com/jnataky/Visual_Analytics/main/Module3/cleaned-cdc-mortality-1999-2010-2.csv")
```


## Question 1

As a researcher, you frequently compare mortality rates from particular causes across different States. You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). Create a visualization that allows you to rank States by crude mortality for each cause of death.


**shinyApp link:**

```{r eval = FALSE}
cdc_data <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv")

library(shiny)
library(shinyWidgets)

# Define UI for application that draws a bar graph


ui <- fluidPage(
    
    # use a gradient in background
    setBackgroundColor(
        color = c("#F7FBFF", "#2171B5"),
        gradient = "linear",
        direction = "bottom"
    ),
    
    # Application title
    titlePanel(div("Cause of Crude Mortality Rate Across US States", style = "color: #ff0000")), 
    
    
    # Sidebar with a slider input for number of bins 
    sidebarLayout(
        selectInput("select", label = h3("Causes of Death"), 
                    choices = cdc_data$ICD.Chapter, 
                    selected = 1,
                    width = '100%'),
        
        # Show a plot of the generated distribution
        
        mainPanel(
            plotOutput("distPlot")
        )
    )
)

# Define server logic required to draw a bar graph

server <- function(input, output) {
    output$distPlot <- renderPlot({
        ggplot(cdc_data[cdc_data$ICD.Chapter == input$select,] , aes(x = reorder(State, Crude.Rate), y = Crude.Rate)) +
            labs(x = "State", y = "Crude Mortality Rate") +  
            geom_bar(stat = "identity", fill = "#ff0000") +
            coord_flip() +
            theme_minimal()
    }, width = 'auto', height = 'auto')
}

# Run the application
shinyApp(ui = ui, server = server)
```



# Question 2

Often you are asked whether particular States are improving their mortality rates (per cause) faster than, or slower than, the national average. Create a visualization that lets your clients see this for themselves for one cause of death at the time. Keep in mind that the national average should be weighted by the national population.

**shinyApp link:**

```{r eval = FALSE}

library(shiny)
library(plotly)
library(rsconnect)

# Import the Data

cdc_data<- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv")


# Define UI for application

ui <- fluidPage(
    
    # use a gradient in background
    setBackgroundColor(
        color = c("#F7FBFF", "#2171B5"),
        gradient = "linear",
        direction = "bottom"
    ),
    
    titlePanel("Crude Mortality Rate Across US States Vs National Average"), 
    sidebarPanel(
        selectInput("select1", label = strong("State"), 
                    choices = levels(as.factor(cdc_data$State)), 
                    selected = 1),
        
        selectInput("select2", label = strong("Cause of Death"), 
                    choices = levels(as.factor(cdc_data$ICD.Chapter)), 
                    selected = 1),
        
        width = "auto"
    ),
    
    mainPanel(
        plotOutput("distPlot")
    )
)


# Define server logic required 

server <- function(input, output) {
    output$distPlot <- renderPlot({
        cdc_data %>% 
            group_by(Year, ICD.Chapter) %>%
            mutate(N_Population = sum(Population),
                   N_Count = sum(Deaths), 
                   N_Crude_Rate = 10^5*(N_Count/N_Population)) %>% 
            group_by(Year, ICD.Chapter, State) %>%
            mutate(S_Count=sum(Deaths),
                   S_Crude_Rate=10^5*(S_Count/Population)) %>%
            
            select(ICD.Chapter, State, Year, N_Crude_Rate, S_Crude_Rate) %>% 
            filter(ICD.Chapter == input$select2, State == input$select1) %>% 
            
            ggplot() +
            geom_bar(aes(x = Year, weight = S_Crude_Rate), fill = "#ff0000") +
            labs(x = "State", y = "Crude Mortality Rate") +
            
            geom_line(aes(x = Year, y = N_Crude_Rate, linetype = "National Average"), col = "blue", lwd = 1) +
            scale_linetype(name = NULL) +
            theme_minimal()
    }
    )
}


# Run the application 
shinyApp(ui = ui, server = server)
```

